---
title: "graphs"
author: "Spencer Rhea"
date: "5/31/2020"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(feather)
library(tidyverse)
library(stringr)
library(viridis)
library(sf)
library(FedData)
library(mapview)
library(lubridate)
library(feather)
library(tidycensus)
library(USAboundaries)
library(ggthemes)
library(ggplot2)
library(kableExtra)
library(knitr)
library(tmap)
library(tmaptools)
library(RColorBrewer)
library(nhdplusTools)
```


# Load and download data
## NHD
```{r, eval=FALSE}
#co_extent <- extent(us_states(states = "CO"))

#co_bbox <- bbox(co_extent)

#subset_nhdplus(
#comids = NULL,
#output_file = "data/nhd/NHDPlus/download.gpkg",
#nhdplus_data = "download",
#bbox = co_bbox,
#simplified = FALSE,
#overwrite = FALSE,
#return_data = TRUE,
#status = TRUE
#)

nhd <- st_read('/Users/spencerrhea/Google Drive/EJ_paper /data/nhd/NHDPlus/download.gpkg', "NHDFlowline_Network")

nhd_bodies <- st_read('/Users/spencerrhea/Google Drive/EJ_paper /data/nhd/NHDPlus/download.gpkg', "NHDWaterbody")
```

## E.coli data 
```{r, eval=FALSE}
# Read in raw e.coli data
ecoli_raw <- read_feather("/Users/spencerrhea/Google Drive/EJ_paper /data/wq_data/Colorado_ecoli.feather")

# Read in inventory
inv <- read_feather('/Users/spencerrhea/Google Drive/EJ_paper /data/wq_data/test/wqp_inv.feather') %>%
    select(SiteID=MonitoringLocationIdentifier,
           resultCount,
           parameter,
                  lat,
                  lon) %>%
  distinct() %>%
  filter(!is.na(lat),
         !is.na(lon)) %>%
  st_as_sf(.,coords=c('lon','lat'),crs=4326) %>%
  distinct(SiteID)

# clean e.coli data
clean_ecoli <-  ecoli_raw %>%
                  dplyr::select(date=ActivityStartDate,
                         parameter=CharacteristicName,
                         units=ResultMeasure.MeasureUnitCode,
                         SiteID=MonitoringLocationIdentifier,
                         org=OrganizationFormalName,
                         org_id=OrganizationIdentifier,
                         time=ActivityStartTime.Time,
                         raw_value=ResultMeasureValue,
                         sample_method=SampleCollectionMethod.MethodName,
                         analytical_method=ResultAnalyticalMethod.MethodName,
                         particle_size=ResultParticleSizeBasisText,
                         date_time=ActivityStartDateTime,
                         media=ActivityMediaName,
                         sample_depth=ActivityDepthHeightMeasure.MeasureValue,
                         sample_depth_unit=ActivityDepthHeightMeasure.MeasureUnitCode,
                         fraction=ResultSampleFractionText,
                         status=ResultStatusIdentifier) %>%
  mutate(units = trimws(units)) %>%
  #Keep only samples that are water samples
  filter(media=='Water') %>%
  filter(units != "ug/l")

# Removed filter for log mean 
ecoli_mean <- clean_ecoli %>%
  filter(year(date) >= 2005 & year(date) <= 2015) %>%
  filter(parameter == "Escherichia coli") %>%
  group_by(SiteID) %>%
  summarize(mean = mean(raw_value,na.rm=T) %>% round(.,2),
            raw_mean = mean(raw_value,na.rm=T),
            sd = sd(raw_value, na.rm = T),
            count=n()) %>%
  filter(count > 9) %>% 
  #filter(sd < 5) %>%
  mutate(log_mean=log10(mean)) %>%
  filter(log_mean > 0.0000001)

ecoli_mean_sf <- inv %>%
  right_join(.,ecoli_mean,by='SiteID') %>%
  st_transform(2163) 

ecoli_mean_sf_100m <- ecoli_mean_sf %>%
  st_buffer(100)

ecoli_3km <- ecoli_mean_sf %>%
  st_buffer(3000)
```


## WWTP 
```{r, eval=FALSE}
#Read in data from EPA's facility registration service, downloaded from EPA website 
wastewater <- st_read(dsn = "/Users/spencerrhea/Google Drive/EJ_paper /data/wastewater/CWA_summaries_060314.gdb") %>%
  filter(CWP_STATE == "CO")


waste_water_full <- wastewater %>%
  st_as_sf(coords = c("coords.x1", "coords.x2"), crs = "+proj=merc +a=6378137 +b=6378137 +lat_ts=0.0 +lon_0=0.0 +x_0=0.0 +y_0=0 +k=1.0 +units=m +nadgrids=@null +no_defs") %>%
  filter(CWP_STATE == "CO") 

#filter to only WWTP and remove permits that are no longer active
sewage <- waste_water_full %>%
  filter(SIC_CODES == 4952) %>%
  filter(CWP_PERMIT_STATUS_DESC != "Terminated") %>%
  mutate(type = "plant") %>%
  filter(CWP_PERMIT_STATUS_DESC != "Expired") %>%
  st_transform(2163) 

#Flow 


#wwtp_flow <- read_csv("/Users/spencerrhea/Google Drive/EJ_paper /data/wwtp_flow.csv") 

#wwtp_flow <- wwtp_flow %>%
#  filter(State == "CO")

#wwtp_loc <- read_rds("/Users/spencerrhea/Google Drive/EJ_paper /data/sewage_plants.rds")
```

## Landcover 
```{r, eval=FALSE}
#Read in landcover database 
nlcd <- raster("/Users/spencerrhea/Google Drive/EJ_paper /data/nlcd/NLCD_2016.img") 

#get colorado shape 
co <- us_states(states = "CO") %>%
  st_transform(5070)

#crop nlcd to colorado
nlcd_co <- crop(nlcd, co)

nlcd_object <- nlcd_co

#get legend for nlcd codes 
codes <- nlcd_co@data@attributes[[1]][["NLCD.Land.Cover.Class"]]

legend <- as_data_frame(codes) %>%
  mutate(id = row_number()) %>%
  rename(type = value)  %>%
  mutate(land = ifelse(type == "", NA, type)) %>%
  filter(!is.na(land)) %>%
  select(-land) %>%
  mutate(type = as.character(type)) %>%
  mutate(id = id-1)
```

## Census data
```{r, message=FALSE, warning=FALSE, eval=FALSE}
#Table to find census parameters of interest 
v12 <- load_variables(2013, "acs5", cache = TRUE)

front_range_co <- c("Adams", "Arapahoe", "Broomfield", "Clear Creek", "Douglas", "El Paso", "Elbert", "Fremont", "Gilpin", "Jefferson", "Larimer", "Park", "Pueblo", "Teller", "Denver")

codes <- c(population = "B01003_001",
           median_income = "B19013_001",
           gini_index = "B19083_001",
           median_age = "B01002_001",
           education_25_total = "B15003_001",
           hs_diploma = "B15003_017",
           GED = "B15003_018",
           some_college_1year = "B15003_019",
           some_college = "B15003_020",
           associates = "B15003_021",
           bachelors = "B15003_022",
           masters = "B15003_023",
           professional = "B15003_024",
           doctoral = "B15003_025",
           tenure_total = "B25003_001",
           renters = "B25003_003",
           owners = "B25003_002",
           sex_total = "B01001_001",
           male = "B01001_002",
           female = "B01001_026",
           race_total = "B03002_001",
           white = "B03002_003",
           black = "B03002_004",
           native = "B03002_005",
           asian = "B03002_006",
           hawaiian = "B03002_007",
           other = "B03002_008",
           two_or_more = "B03002_009",
           hispanic = "B03002_012")

#Download block group geometries from tidycensus package 
block_group_geom <- get_acs(geography = "block group", variables = "B01003_001", year = 2013, state = "CO", county = front_range_co, geometry = T) %>%
  select(GEOID, geometry)

#Download block group census data
block_group_census <- get_acs(geography = "block group", codes, year = 2013, state = "CO", county = front_range_co, geometry = F) 

#Convert census data into wide table and calculate parameters of interest 
bg_wide <- block_group_census %>%
  pivot_wider(names_from = c("variable"), values_from = c("estimate", "moe")) %>%
  mutate(color = estimate_black+estimate_native+estimate_asian+estimate_hawaiian+estimate_other+estimate_two_or_more+estimate_hispanic,
         degrees = estimate_associates+estimate_bachelors+estimate_masters+estimate_professional+estimate_doctoral) %>%
  rename(renters = estimate_renters,
         female = estimate_female,
         population = estimate_population,
         renters = estimate_renters,
         tenure_total = estimate_tenure_total,
         education_25_total = estimate_education_25_total,
         median_income = estimate_median_income,
         median_age = estimate_median_age)

#Join data with geometries
bg_wide_geom <- inner_join(block_group_geom, bg_wide, by = "GEOID") %>%
  st_transform(2163)

#write_rds(bg_wide_geom, "data/census_fr.rds")
```


# Data Avilible 
### Only front range counties 
```{r, eval=FALSE}
fr_census <- read_rds("/Users/spencerrhea/Google Drive/EJ_paper /data/census_fr.rds")

nhd_order <- nhd %>%
  st_transform(2163) %>%
  filter(streamorde > 2) 

nhd_bodies_trans <- nhd_bodies %>%
  filter(ftype != "Ice Mass") %>%
  st_transform(2163)

#Gen number of river segments in block group
data_avail <- fr_census %>%
  st_buffer(3000) %>%
  st_join(nhd_order) %>%
  group_by(GEOID) %>%
  summarise(num = n(),
            length = sum(lengthkm, na.rm = T)) %>%
  mutate(num = ifelse(length <= 0.001, 0, num))

#number of lakes near block groups 
data_bodies <- fr_census %>%
  st_buffer(3000) %>%
  st_join(nhd_bodies_trans) %>%
  group_by(GEOID) %>%
  summarise(lakes = n(),
            area = sum(areasqkm, na.rm = T)) %>%
  mutate(lakes = ifelse(area <= 0.0001, 0, lakes))

#combine lakes and streams 
data_full <- data_avail %>%
  as.data.frame() %>%
  select(-geometry) %>%
  full_join(as.data.frame(data_bodies), by = "GEOID") %>%
  select(-geometry) %>%
  mutate(water = ifelse(lakes == 0 & num == 0, "non", "yes"))

#join nhd to census 
ecoli_avail <- st_join(fr_census, ecoli_3km) %>%
  mutate(data_avail = ifelse(!is.na(mean),'has_data','no_data')) %>%
  mutate(data_avail = as.factor(data_avail)) %>%
  as.data.frame() %>%
  select(-geometry)

ecoli_water <- full_join(data_full, ecoli_avail, by = "GEOID") %>%
  mutate(percent_color = color/population,
         percent_renters = renters/tenure_total,
         percent_degrees = degrees/education_25_total) 

#write_rds(ecoli_water, "data_out/data_avail.rds")

ecoli_water_graphe <- ecoli_water %>%
  select(water, POC=percent_color, renters=percent_renters, degrees=percent_degrees, income=median_income, water) %>%
  pivot_longer(cols = c(POC, renters, degrees, income))
```

## Lives within 3 km of a waterbody 
```{r, eval=FALSE}
ggplot(ecoli_water_graphe, aes(x = water, y = value)) +
  geom_violin(draw_quantiles = 0.5) +
  theme_few() +
  facet_wrap(~name, scales = "free") +
  labs(x = "3km within a waterbody", y = "Proportion people of color")
```

## Of areas within 3 km of a waterbody, ecoli availability 
```{r, eval=F}
ecoli_water_sub <- ecoli_water %>%
  filter(water == "yes") %>%
  select(data_avail, POC=percent_color, renters=percent_renters, degrees=percent_degrees, income=median_income, water) %>%
  pivot_longer(cols = c(POC, renters, degrees, income))

ggplot(ecoli_water_sub, aes(x = data_avail, y = value)) +geom_violin(draw_quantiles = 0.5) +
  theme_few() +
  facet_wrap(~name, scales = "free") +
  labs(x = "E.coli data availability", y = "Proportion people of color")
```

# Function 
### Downloads each river network and gets the percent of landcover at each river segment on the main stem. Also finds location of WWTP on the main stem. 
```{r, eval=FALSE}
#This function joins all sub basins in a basin to the nlcd and gets the number of pixles for all catagories 
define_land_by_catch <- function(catch, row) {

#transform to the nlcd crs 
whole_catch <- catch[row,] %>%
  st_transform(5070)

#pull the basin codes
fetid <- pull(whole_catch, featureid)

#get etent of basin
catch_extent <- extent(whole_catch)

#crops nlcd to basin 
nlcd_crop <- crop(nlcd_object, catch_extent) 

nlcd_mask <- mask(nlcd_crop, whole_catch) 

nlcd_trim <- trim(nlcd_mask) 

#gets the numver of pixles in sub basin 
values <- as.data.frame(getValues(nlcd_trim)) %>%
  rename(id = 1) %>%
  group_by(id) %>%
  summarise(num = n()) %>%
  filter(id != 0) %>%
  full_join(legend) %>%
  filter(!is.na(id))%>%
  mutate(featureid = fetid) %>%
  select(-id) 

#Gets the total number of pixles in a basin
total <- sum(values$num, na.rm = T)

#pivot the table to wide
fin <- values %>%
  pivot_wider(names_from = type, values_from = num) %>%
  mutate(Total = total)
}



#This function calculates the percent of landocover for all stream segmants on the main stem of the river 
percecnt_along <- function(river_comid, flow_lines = flowline, watersheds = landcover_by_catchment) {

#gets all sub basins that contribute to a river segmant 
up_stream <- get_UT(flow_lines, river_comid)

#filter flowlines to only ones contributing to a point on the main stem 
up_segments <- flow_lines %>%
  filter(comid %in% up_stream)

#filter catchment to only the basin contributing to a point on the main stem 
catch <- watersheds %>%
  filter(featureid %in% pull(up_segments, comid)) %>%
  pivot_longer(cols = c("Evergreen Forest", "Shrub/Scrub", "Unclassified", "Open Water", "Perennial Snow/Ice", "Developed, Open Space", "Developed, Low Intensity", "Developed, Medium Intensity", "Developed, High Intensity", "Barren Land", "Deciduous Forest", "Mixed Forest", "Herbaceuous", "Hay/Pasture", "Cultivated Crops", "Woody Wetlands", "Emergent Herbaceuous Wetlands")) %>%
  group_by(name) %>%
  summarise(num = sum(value, na.rm = T)) %>%
  mutate(num = as.numeric(num)) 

#gets total number of pixles for each catchment 
total <- sum(catch$num) 

#gets percent of landcover for each catchment 
final <- catch %>%
  mutate(Total = total) %>%
  mutate(percent = (num/Total)*100) %>%
  mutate(percent = round(percent, 5)) %>%
  select(name, percent) %>%
  mutate(comid = river_comid) %>%
  pivot_wider(names_from = name, values_from = percent)
}



#This function return the landcover percentages for all subasins 
river_longitudinal <- function(lat, long) {

start <- st_sfc(st_point(c(long, lat)), crs = 4269)

#Gets river segment of where the point is 
start_comid <- discover_nhdplus_id(start)

#get flowlines for river network
river_flowline <- navigate_nldi(list(featureSource = "comid", 
                               featureID = start_comid), 
                          mode = "upstreamTributaries", 
                          data_source = "")

#get catchments, waterbodies, and flowline 
river_subset <- subset_nhdplus(comids = river_flowline$nhdplus_comid,
                         output_file = "data/nhd/subset.gpkg",
                         nhdplus_data = "download", 
                         return_data = TRUE, 
                         overwrite = T)

#Pull out all dataframes 
river_flowline <- river_subset$NHDFlowline_Network
river_catchment <- river_subset$CatchmentSP
river_waterbody <- river_subset$NHDWaterbody

#prep dataframe for nhdpulstools 
river_prep <- prepare_nhdplus(river_flowline, min_network_size = 10, min_path_length = 10) %>%
  select(ID = COMID, toID = toCOMID, length = LENGTHKM)

#get the mainstem flowline codes
river_main <- get_UM(river_flowline, start_comid)

#get dataframe with only mainstem 
river <- river_flowline %>%
  filter(comid %in% river_main)

#Get pathlength of river segments 
river_pathlength <- get_pathlength(river_prep)

#join pathlengths with dataframe 
river_length <- river_flowline %>%
  rename(ID = comid) %>%
  right_join(river_pathlength, by = "ID")

#join mainstem to pethlengths 
riv <- river_length %>%
  filter(ID %in% river_main) %>%
  st_transform(4269) %>%
  select(-pathlength.x) %>%
  rename(pathlength = pathlength.y)

#add id column 
river_catchment_id <- river_catchment %>%
  mutate(row = row_number())

whole_catch <- river_catchment_id

# Us function to get landcover for all catchments 
river_landcover_by_catchment <- define_land_by_catch(river_catchment_id, 1)

land_length <- as.numeric(length(river_catchment_id$id))

seqens <- 1:(land_length-1)
for(i in seqens) {
  x <- i+1
  shed <- define_land_by_catch(river_catchment_id, x)
  
  river_landcover_by_catchment <- rbind(river_landcover_by_catchment, shed)
}

landcover_by_catchment <- river_landcover_by_catchment

flowline <- river_flowline

#get percentage for all river segments on the mainstem 
river_comids <- pull(riv, ID)

river_percents <- percecnt_along(river_comids[1], flowline, landcover_by_catchment)

comids_length <- as.numeric(length(river_comids)[1])

seq <- 1:(comids_length-1)
for(i in seq) {
  new <- percecnt_along(river_comids[i+1], flowline, landcover_by_catchment) 
  
  river_percents <- rbind(new, river_percents)
}

#define all landcover types 

river_percent_along_river <- river_percents %>%
  rename(barren = 2,
         crops = 3,
         deciduous = 4,
         developed_high = 5,
         developed_low = 6,
         developed_mid = 7,
         developed_open = 8,
         wetland_herbaceous = 9,
         evergreen = 10,
         hay_pasture = 11,
         herbaceous = 12,
         forest_mixed = 13,
         water = 14,
         ice = 15,
         shrub = 16,
         wetland_wood = 18) %>%
  select(-Unclassified)


#summarise landcover types 
river_landcover_sumz <- river_percent_along_river %>%
  mutate(developed = developed_high+developed_low+developed_mid+developed_open,
         forest = forest_mixed+deciduous+evergreen,
         wetland = wetland_herbaceous+wetland_wood,
         agriculture = crops+hay_pasture) %>%
  select(-developed_high, -developed_mid, -developed_low, -forest_mixed, -deciduous, -evergreen, -wetland_herbaceous, -wetland_wood, -crops, -hay_pasture, -developed_open) 

#get laength of river 
max_length <- max(riv$pathlength)

#change length so the outlet is the largest number 
riv_trans <- riv %>%
  mutate(pathlength = max_length-pathlength) %>%
  rename(comid = ID) %>%
  st_transform(2163) 

#buffer sewage to join to river 
wwtp_buff <- sewage %>%
  st_buffer(300)

#get all wwtp on the river 
river_wwtp <- st_join(riv_trans, wwtp_buff) %>%
  filter(!is.na(CWP_NAME))

#get only codes 
river_sweage_codes <- river_wwtp %>%
  group_by(REGISTRY_ID) %>%
  summarise(length = mean(pathlength, na.rm = T)) %>%
  pull(length)

river_sweage_id <- river_wwtp %>%
  group_by(REGISTRY_ID) %>%
  summarise(length = mean(pathlength, na.rm = T)) %>%
  pull(REGISTRY_ID)

river_sweage_names <- river_wwtp %>%
  filter(REGISTRY_ID %in% river_sweage_id) %>%
  pull(CWP_NAME)

#join ecoli points to rivers
riv_jumps <- st_join(riv_trans, ecoli_mean_sf_100m)

#make dataframe with landcover and ecoli points 
riv_jumps_fil <- riv_jumps %>%
  as.data.frame() %>%
  full_join(river_landcover_sumz, by = "comid") %>%
  select(-geometry) %>%
  filter(!is.na(pathlength)) %>%
  select(pathlength, mean, herbaceous, shrub, developed, forest, wetland, agriculture) %>%
  mutate(mean = mean/50) %>%
   pivot_longer(cols = c(herbaceous, shrub, developed, forest, wetland, agriculture))

#make a list with both landcover dataframe and wwtp locations 
fin <- list(full = riv_jumps_fil,
     wwtp = river_sweage_codes,
     wwtp_names = river_sweage_names,
     nlcd_raw = river_percent_along_river,
     flowlines = river_flowline, 
     catchment = river_catchment,
     waterbodies = river_waterbody, 
     mainstem = riv)

return(fin)

}
```

# Stream Jumps 
### Vertical lines represent WWTP, some lines close together are probably a double count 
## Poudre 
```{r, warning=FALSE}
# poudre_test <- river_longitudinal(lat = 40.423795, long = -104.600462)
# 
# write_rds(poudre_test, "data_out/poudre.rds")

poudre <- read_rds("/Volumes/Spencer backup/Google Drive/EJ_paper /data_out/poudre.rds")

ggplot(poudre[["full"]], aes(x = pathlength, y = value, colour = name)) +
  geom_vline(xintercept = poudre[["wwtp"]], color = "gray") +
  geom_line() +
  geom_point(aes(x = pathlength, y = mean)) +
  scale_color_manual(values = c("#252525", "#e7298a", "#66a61e", "#7570b3", "#d95f02", "#1b9e77"), 
                     breaks = c("herbaceous", "shrub", "forest", "wetland", "agriculture", "developed"),
                     name = "Landcover") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*50,  name = "E.coli (#/ml)")) +
  theme_few() +
  labs(x = "Length on River", y = "Watershed landcover (%)")
```

## Big Thompson 
```{r, warning=FALSE, echo=FALSE}
# bt_full <- river_longitudinal(40.354054, -104.761314)
# 
# write_rds(bt_full, "data_out/big_thompson.rds")

bt_full <- read_rds("/Volumes/Spencer backup/Google Drive/EJ_paper /data_out/big_thompson.rds")

ggplot(bt_full[["full"]], aes(x = pathlength, y = value, colour = name)) +
  geom_vline(xintercept = bt_full[["wwtp"]], color = "gray") +
  geom_line() +
  geom_point(aes(x = pathlength, y = mean)) +
  scale_color_manual(values = c("#252525", "#e7298a", "#66a61e", "#7570b3", "#d95f02", "#1b9e77"), 
                     breaks = c("herbaceous", "shrub", "forest", "wetland", "agriculture", "developed"),
                     name = "Landcover") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*50,  name = "E.coli (#/ml)")) +
  theme_few() +
  labs(x = "Length on River", y = "Watershed landcover (%)")
```

## St. Vrain 
```{r, warning=FALSE, echo=FALSE}
# st_vrain <- river_longitudinal(40.270375, -104.874370)
# 
# write_rds(st_vrain, "data_out/st_vrain.rds")

st_vrain <- read_rds("/Volumes/Spencer backup/Google Drive/EJ_paper /data_out/st_vrain.rds")

ggplot(st_vrain[["full"]], aes(x = pathlength, y = value, colour = name)) +
  geom_vline(xintercept = st_vrain[["wwtp"]], color = "gray") +
  geom_line() +
  geom_point(aes(x = pathlength, y = mean)) +
  scale_color_manual(values = c("#252525", "#e7298a", "#66a61e", "#7570b3", "#d95f02", "#1b9e77"), 
                     breaks = c("herbaceous", "shrub", "forest", "wetland", "agriculture", "developed"),
                     name = "Landcover") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*50,  name = "E.coli (#/ml)")) +
  theme_few() +
  labs(x = "Length on River", y = "Watershed landcover (%)")
```

## Boulder Creek 
```{r, warning=FALSE, echo=FALSE}
# boulder_creek <- river_longitudinal(40.157845, -105.009788)
# 
# write_rds(boulder_creek, "data_out/boulder_creek.rds")

boulder_creek <- read_rds("/Volumes/Spencer backup/Google Drive/EJ_paper /data_out/boulder_creek.rds")

ggplot(boulder_creek[["full"]], aes(x = pathlength, y = value, colour = name)) +
  geom_vline(xintercept = boulder_creek[["wwtp"]], color = "gray") +
  geom_line() +
  geom_point(aes(x = pathlength, y = mean)) +
  scale_color_manual(values = c("#252525", "#e7298a", "#66a61e", "#7570b3", "#d95f02", "#1b9e77"), 
                     breaks = c("herbaceous", "shrub", "forest", "wetland", "agriculture", "developed"),
                     name = "Landcover") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*50,  name = "E.coli (#/ml)")) +
  theme_few() +
  labs(x = "Length on River", y = "Watershed landcover (%)")
```


## South Platte 
```{r, warning=FALSE, echo=FALSE}
# south_platte <- river_longitudinal(40.629481, -103.181869)
# 
# write_rds(south_platte, "data_out/south_platte.rds")

south_platte <- read_rds("/Volumes/Spencer backup/Google Drive/EJ_paper /data_out/south_platte.rds")

sp_sweage_codes <- c(281.122, 281.122, 188.124, 188.124, 178.222, 178.222, 178.222, 178.222, 178.222, 178.222, 157.964, 157.964, 157.964, 157.964, 157.964, 157.964)

ggplot(south_platte, aes(x = pathlength, y = value, colour = name)) +
  geom_vline(xintercept = sp_sweage_codes, color = "gray") +
  geom_line() +
  geom_point(aes(x = pathlength, y = mean)) +
  scale_color_manual(values = c("#252525", "#e7298a", "#66a61e", "#7570b3", "#d95f02", "#1b9e77"), 
                     breaks = c("herbaceous", "shrub", "forest", "wetland", "agriculture", "developed"),
                     name = "Landcover") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*50,  name = "E.coli (#/ml)")) +
  theme_few() +
  labs(x = "Length on River", y = "Watershed landcover (%)")
```

## Monument Creek 
```{r, warning=FALSE, echo=FALSE}
# monument_creek <- river_longitudinal(38.832856, -104.831865)
# 
# write_rds(monument_creek, "data_out/monument_creek.rds")

monument_creek <- read_rds("/Volumes/Spencer backup/Google Drive/EJ_paper /data_out/monument_creek.rds")

ggplot(monument_creek[["full"]], aes(x = pathlength, y = value, colour = name)) +
  geom_vline(xintercept = monument_creek[["wwtp"]], color = "gray") +
  geom_line() +
  geom_point(aes(x = pathlength, y = mean)) +
  scale_color_manual(values = c("#252525", "#e7298a", "#66a61e", "#7570b3", "#d95f02", "#1b9e77"), 
                     breaks = c("herbaceous", "shrub", "forest", "wetland", "agriculture", "developed"),
                     name = "Landcover") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*50,  name = "E.coli (#/ml)")) +
  theme_few() +
  labs(x = "Length on River", y = "Watershed landcover (%)")
```


## Fountain Creek
```{r, warning=FALSE, echo=FALSE}
# fountain_full <- river_longitudinal(38.254237, -104.589736)
# 
# write_rds(fountain_full, "data_out/fountain.rds")

fountain_full <- read_rds("/Volumes/Spencer backup/Google Drive/EJ_paper /data_out/fountain.rds")

ggplot(fountain_full[["full"]], aes(x = pathlength, y = value, colour = name)) +
  geom_vline(xintercept = fountain_full[["wwtp"]], color = "gray") +
  geom_line() +
  geom_point(aes(x = pathlength, y = mean)) +
  scale_color_manual(values = c("#252525", "#e7298a", "#66a61e", "#7570b3", "#d95f02", "#1b9e77"), 
                     breaks = c("herbaceous", "shrub", "forest", "wetland", "agriculture", "developed"),
                     name = "Landcover") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*50,  name = "E.coli (#/ml)")) +
  theme_few() +
  labs(x = "Length on River", y = "Watershed landcover (%)")
```


## Arkansas 
```{r, warning=FALSE, echo=FALSE}
# arkansas <- river_longitudinal(38.081337, -103.232209)
# 
# write_rds(arkansas, "data_out/arkansas.rds")

arkansas_full <- read_rds("/Volumes/Spencer backup/Google Drive/EJ_paper /data_out/arkansas.rds")

ggplot(arkansas_full[["full"]], aes(x = pathlength, y = value, colour = name)) +
  geom_vline(xintercept = arkansas_full[["wwtp"]], color = "gray") +
  geom_line() +
  geom_point(aes(x = pathlength, y = mean)) +
  scale_color_manual(values = c("#252525", "#e7298a", "#66a61e", "#7570b3", "#d95f02", "#1b9e77"), 
                     breaks = c("herbaceous", "shrub", "forest", "wetland", "agriculture", "developed"),
                     name = "Landcover") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*50,  name = "E.coli (#/ml)")) +
  theme_few() +
  labs(x = "Length on River", y = "Watershed landcover (%)")
```


