---
title: "front_range_model"
author: "Spencer Rhea"
date: "6/5/2020"
output: 
  html_document: 
    toc: true
    toc_depth: 2
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE)
library(raster)
library(feather)
library(tidyverse)
library(stringr)
library(ggpubr)
library(viridis)
library(sf)
library(FedData)
library(mapview)
library(lubridate)
library(feather)
library(tidycensus)
library(USAboundaries)
library(ggthemes)
library(ggplot2)
library(kableExtra)
library(knitr)
library(tmap)
library(tmaptools)
library(RColorBrewer)
library(nhdplusTools)
```

# Intro 
### We used the same general approach but added and fixed a few things to make the analysis stronger. We refined the data/no data to first look at block groups that are within 3 km and then looked to see if these block groups had ecoli data available. We also ran the linear regression model for only the front range (not the whole state) and ran individual models for each city. This allows us to compare cities to one another and to front range as a whole. Finally we created graphs to relate ecoli concentration, landcover, and wastewater treatment plants (WWTP).


# Data Availability 
## Block groups within 3 km of a waterbody 
### These graphs show the median values and distributions of the four demographic parameters we were interested in between block groups that are within 3km of a waterbody (lakes, ponds, and third order streams or greater) and not within 3 km of a waterbody.  


```{r}
ecoli_water <- read_rds("/Users/spencerrhea/Google Drive/EJ_paper /data_out/data_avail.rds")

ecoli_water_graphe <- ecoli_water %>%
  select(water, POC=percent_color, renters=percent_renters, degrees=percent_degrees, income=median_income, water) %>%
  pivot_longer(cols = c(POC, renters, degrees, income))
```

```{r, include=TRUE, warning=FALSE}
ggplot(ecoli_water_graphe, aes(x = water, y = value)) +
  geom_violin(draw_quantiles = 0.5) +
  theme_few() +
  facet_wrap(~name, scales = "free") +
  labs(x = "3km within a waterbody", y = "Proportion or Median Income")
```

### We see here that block groups that are within 3 km of a waterbody/stream have lower levels of education, lower income, more POC, and more renters. On a first glance, this seems to follow some other work that showed in some areas of the US marginalized groups are more exposed to flooding because they live closer to streams or in the floodplain.

## Of block groups within 3 km of a waterbody, ecoli data availability 
```{r}
ecoli_water_sub <- ecoli_water %>%
  filter(water == "yes") %>%
  select(data_avail, POC=percent_color, renters=percent_renters, degrees=percent_degrees, income=median_income, water) %>%
  pivot_longer(cols = c(POC, renters, degrees, income))

```

```{r, include=TRUE, warning=FALSE}
ggplot(ecoli_water_sub, aes(x = data_avail, y = value)) +
  geom_violin(draw_quantiles = 0.5) +
  theme_few() +
  facet_wrap(~name, scales = "free") +
  labs(x = "E.coli data availability", y = "Proportion or Median Income")
```

### Here we see a similar result from the thesis, where the medians values of areas with ecoli data and without are small for POC and education level (Degrees) but larger for percent renters and income.

```{r}
census <- read_rds("/Users/spencerrhea/Google Drive/EJ_paper /data/census_fr.rds")

city_geom <- get_acs(geography = "metropolitan statistical area/micropolitan statistical area", variables = "B01003_001", year = 2013, geometry = T) %>%
  filter(NAME %in% c("Fort Collins, CO Metro Area", "Denver-Aurora-Lakewood, CO Metro Area", "Colorado Springs, CO Metro Area", "Pueblo, CO Metro Area")) %>%
  st_transform(2163)
```


```{r}
ecoli_raw <- read_feather("/Users/spencerrhea/Google Drive/EJ_paper /data/wq_data/Colorado_ecoli.feather")

# Read in inventory
inv <- read_feather('/Users/spencerrhea/Google Drive/EJ_paper /data/wq_data/test/wqp_inv.feather') %>%
    select(SiteID=MonitoringLocationIdentifier,
           resultCount,
           parameter,
                  lat,
                  lon) %>%
  distinct() %>%
  filter(!is.na(lat),
         !is.na(lon)) %>%
  st_as_sf(.,coords=c('lon','lat'),crs=4326) %>%
  distinct(SiteID)

# clean e.coli data
clean_ecoli <-  ecoli_raw %>%
                  dplyr::select(date=ActivityStartDate,
                         parameter=CharacteristicName,
                         units=ResultMeasure.MeasureUnitCode,
                         SiteID=MonitoringLocationIdentifier,
                         org=OrganizationFormalName,
                         org_id=OrganizationIdentifier,
                         time=ActivityStartTime.Time,
                         raw_value=ResultMeasureValue,
                         sample_method=SampleCollectionMethod.MethodName,
                         analytical_method=ResultAnalyticalMethod.MethodName,
                         particle_size=ResultParticleSizeBasisText,
                         date_time=ActivityStartDateTime,
                         media=ActivityMediaName,
                         sample_depth=ActivityDepthHeightMeasure.MeasureValue,
                         sample_depth_unit=ActivityDepthHeightMeasure.MeasureUnitCode,
                         fraction=ResultSampleFractionText,
                         status=ResultStatusIdentifier) %>%
  mutate(units = trimws(units)) %>%
  #Keep only samples that are water samples
  filter(media=='Water') %>%
  filter(units != "ug/l")

# Removed filter for log mean 
ecoli_mean <- clean_ecoli %>%
  filter(year(date) >= 2005 & year(date) <= 2015) %>%
  filter(parameter == "Escherichia coli") %>%
  group_by(SiteID) %>%
  summarize(mean = mean(raw_value,na.rm=T) %>% round(.,2),
            raw_mean = mean(raw_value,na.rm=T),
            sd = sd(raw_value, na.rm = T),
            count=n()) %>%
  filter(count > 9) %>% 
  #filter(sd < 5) %>%
  mutate(log_mean=log10(mean)) %>%
  filter(log_mean > 0.0000001)

ecoli_mean_sf <- inv %>%
  right_join(.,ecoli_mean,by='SiteID') %>%
  st_transform(2163) 

ecoli_3km <- ecoli_mean_sf %>%
  st_buffer(3000)
```


```{r}
colorado_3km_ecoli <- st_join(census, ecoli_3km) %>%
  mutate(data_avail = ifelse(!is.na(mean),'has_data','no_data')) %>%
  mutate(data_avail = as.factor(data_avail)) 

#Removing sites where there is overlap
grouped_3km <- colorado_3km_ecoli %>%
  filter(data_avail == "has_data") %>%
  mutate(weighted_income = median_income*population) %>%
  mutate(weighted_mean = mean*count) %>%
  group_by(SiteID) %>%
  summarise(color = sum(color, na.rm = T),
            mean = mean(mean, na.rm = T),
            weighted_mean = mean(weighted_mean, na.rm = T),
            mean_n = mean(count, na.rm = T),
            total = sum(population, na.rm = T),
            percent_color = (color/total)*100,
            income = sum(weighted_income, na.rm = T)/sum(total),
            renters = sum(renters, na.rm = T),
            tenure_total = sum(tenure_total, na.rm = T),
            degrees = sum(degrees, na.rm = T),
            education = sum(education_25_total, na.rm = T),
            percent_renters = sum(renters, na.rm = T)/sum(tenure_total, na.rm = T),
            percent_degree = sum(degrees, na.rm = T)/sum(education_25_total, na.rm = T),
            age = mean(median_age, na.rm = T),
            gini = mean(estimate_gini_index, na.rm = T),
            percent_black = (sum(estimate_black, na.rm = T)/sum(population, na.rm = T))*100,
            percent_hispanic = (sum(estimate_hispanic, na.rm = T)/sum(population))*100,
            percent_asain = (sum(estimate_asian, na.rm = T)/sum(population, na.rm = T))*100,
            percent_native = (sum(estimate_native, na.rm = T)/sum(population, na.rm = T))*100,
            percent_other_mixed = (sum(estimate_other, na.rm = T)+sum(estimate_two_or_more))/sum(population, na.rm = T)*100,
            percent_hawaiian = sum(estimate_hawaiian, na.rm = T)/sum(population, na.rm = T)*100) %>%
  #Regroup by income
  group_by(income) %>%
  summarise(mean = sum(weighted_mean, na.rm = T)/sum(mean_n, na.rm = T),
            percent_color = sum(color)/sum(total)*100,
            percent_renters = sum(renters, na.rm = T)/sum(tenure_total, na.rm = T)*100,
            percent_degrees = sum(degrees, na.rm = T)/sum(education, na.rm = T)*100) 

grouped_3km_lm <- grouped_3km %>%
  mutate(log_color = log(percent_color+1),
         log_renters = log(percent_renters+1),
         log_degrees = log(percent_degrees+1),
         log_mean = log(mean),
         log_income = log(income)) %>%
  select(log_mean, log_color, log_renters, log_degrees, log_income)

summary(lm(log_mean ~ log_color+log_renters+log_income+log_degrees, grouped_3km_lm))

all_global <- lm(log_mean ~ log_color+log_renters+log_income+log_degrees, grouped_3km_lm)

front_range_m <- tibble("Inputs Parameters" = c("Intercept", "Log Percent People of Color", "Log Percent Renters", "Log Median Income", "Log Percent Collage Degrees"),
       "Estimates" = c(-0.5394, 1.2832, 0.6151, -0.3234, 0.8651),
       "Std. Error" = c(4.8592, 0.2332, 0.2362, 0.5055, 0.4757),
       "p-value" = c(0.9117, 0.00005, 0.0101, 0.5232, 0.0709),
       "Location" = "FR")
```

```{r}
census_ecoli <- function(df_census, ecoli) {
colorado_3km_ecoli <- st_join(df_census, ecoli) %>%
  mutate(data_avail = ifelse(!is.na(mean),'has_data','no_data')) %>%
  mutate(data_avail = as.factor(data_avail)) 

#Removing sites where there is overlap
grouped_3km <- colorado_3km_ecoli %>%
  filter(data_avail == "has_data") %>%
  mutate(weighted_income = median_income*population) %>%
  mutate(weighted_mean = mean*count) %>%
  group_by(SiteID) %>%
  summarise(color = sum(color, na.rm = T),
            mean = mean(mean, na.rm = T),
            weighted_mean = mean(weighted_mean, na.rm = T),
            mean_n = mean(count, na.rm = T),
            total = sum(population, na.rm = T),
            percent_color = (color/total)*100,
            income = sum(weighted_income, na.rm = T)/sum(total),
            renters = sum(renters, na.rm = T),
            tenure_total = sum(tenure_total, na.rm = T),
            degrees = sum(degrees, na.rm = T),
            education = sum(education_25_total, na.rm = T),
            percent_renters = sum(renters, na.rm = T)/sum(tenure_total, na.rm = T),
            percent_degree = sum(degrees, na.rm = T)/sum(education_25_total, na.rm = T),
            age = mean(median_age, na.rm = T),
            gini = mean(estimate_gini_index, na.rm = T),
            percent_black = (sum(estimate_black, na.rm = T)/sum(population, na.rm = T))*100,
            percent_hispanic = (sum(estimate_hispanic, na.rm = T)/sum(population))*100,
            percent_asain = (sum(estimate_asian, na.rm = T)/sum(population, na.rm = T))*100,
            percent_native = (sum(estimate_native, na.rm = T)/sum(population, na.rm = T))*100,
            percent_other_mixed = (sum(estimate_other, na.rm = T)+sum(estimate_two_or_more))/sum(population, na.rm = T)*100,
            percent_hawaiian = sum(estimate_hawaiian, na.rm = T)/sum(population, na.rm = T)*100) %>%
  #Regroup by income
  group_by(income) %>%
  summarise(mean = sum(weighted_mean, na.rm = T)/sum(mean_n, na.rm = T),
            percent_color = sum(color)/sum(total)*100,
            percent_renters = sum(renters, na.rm = T)/sum(tenure_total, na.rm = T)*100,
            percent_degrees = sum(degrees, na.rm = T)/sum(education, na.rm = T)*100) 

grouped_3km_lm <- grouped_3km %>%
  mutate(log_color = log(percent_color+1),
         log_renters = log(percent_renters+1),
         log_degrees = log(percent_degrees+1),
         log_mean = log(mean),
         log_income = log(income)) %>%
  select(log_mean, log_color, log_renters, log_degrees, log_income)

return(grouped_3km_lm) }
```

```{r}
ft_ecoli <- st_filter(ecoli_3km, filter(city_geom, NAME == "Fort Collins, CO Metro Area"))

ft_joined <- census_ecoli(census, ft_ecoli)

summary(lm(log_mean ~ log_color+log_renters+log_income+log_degrees, ft_joined))

ft_m <- tibble("Inputs Parameters" = c("Intercept", "Log Percent People of Color", "Log Percent Renters", "Log Median Income", "Log Percent Collage Degrees"),
       "Estimates" = c(-58.7324, -0.9195, 0.9231, 6.7570 , -3.3689),
       "Std. Error" = c(47.6229, 1.1567, 1.1533,5.8048, 5.3856),
       "pvalue" = c(0.241,0.442,0.439,0.267,0.543),
       "Location" = "FC") %>%
  pivot_longer(cols = c("Estimates", "Std. Error"))
```


```{r}
denver_ecoli <- st_filter(ecoli_3km, filter(city_geom, NAME == "Denver-Aurora-Lakewood, CO Metro Area"))

denver_joined <- census_ecoli(census, denver_ecoli)

summary(lm(log_mean ~ log_color+log_renters+log_income+log_degrees, denver_joined))

denver_m <- tibble("Inputs Parameters" = c("Intercept", "Log Percent People of Color", "Log Percent Renters", "Log Median Income", "Log Percent Collage Degrees"),
       "Estimates" = c(1.428, 2.0203, -0.2857, -0.5874, 1.1359),
       "Std. Error" = c(4.5728, 0.2205, 0.2068,0.5343, 0.4857),
       "pvalue" = c(0.7675,0.000000000005,0.173,0.2766,0.0232),
       "Location" = "Denver") %>%
  pivot_longer(cols = c("Estimates", "Std. Error"))
```

```{r}
csprings_ecoli <- st_filter(ecoli_3km, filter(city_geom, NAME == "Colorado Springs, CO Metro Area"))

csprings_joined <- census_ecoli(census, csprings_ecoli)

summary(lm(log_mean ~ log_color+log_renters+log_income+log_degrees, csprings_joined))

csprings_m <- tibble("Inputs Parameters" = c("Intercept", "Log Percent People of Color", "Log Percent Renters", "Log Median Income", "Log Percent Collage Degrees"),
       "Estimates" = c(-50.3759, 0.8685, 3.2837, 4.0850, -0.6042),
       "Std. Error" = c(16.6753, 0.5805, 1.0540,1.7162, 1.5326),
       "pvalue" = c(0.00462,0.14333,0.00360,0.02271,0.69574),
       "Location" = "CSprings") %>%
  pivot_longer(cols = c("Estimates", "Std. Error"))
```

```{r}
pueblo_ecoli <- st_filter(ecoli_3km, filter(city_geom, NAME == "Pueblo, CO Metro Area"))

pueblo_joined <- census_ecoli(census, pueblo_ecoli)

summary(lm(log_mean ~ log_color+log_renters+log_income+log_degrees, pueblo_joined))

pueblo_m <- tibble("Inputs Parameters" = c("Intercept", "Log Percent People of Color", "Log Percent Renters", "Log Median Income", "Log Percent Collage Degrees"),
       "Estimates" = c(30.157, 5.762, -4.369, -4.738, 5.607),
       "Std. Error" = c(28.246, 2.010, 1.856,3.237, 3.316),
       "pvalue" = c(0.3051,0.0132,0.035,0.167,0.1146),
       "Location" = "Pueblo") %>%
  pivot_longer(cols = c("Estimates", "Std. Error"))
```

```{r}
all_cities <- rbind(pueblo_m, csprings_m, denver_m, ft_m) 

box_color <- all_cities %>%
  filter(name == "Estimates") %>%
  rename(parameter = `Inputs Parameters`) %>%
  mutate(Significance = ifelse(pvalue <= 0.05, "< 0.05", "not")) %>%
  filter(parameter == "Log Percent People of Color") %>%
  filter(Significance == "< 0.05")

color <- ggplot(box_color, aes(x = Location, y = value, colour = Significance, shape = Significance, fill = Significance)) +
  geom_point() +
  theme_few() +
  scale_shape_manual(values = c(15,16),
                     breaks = c("< 0.05", "not"),
                      labels = c("< 0.05", "not")) +
  scale_colour_manual(values = c("#e41a1c", "#4daf4a"),
                      breaks = c("< 0.05", "not"),
                      labels = c("< 0.05", "not")) +
  scale_y_continuous(limits = c(-5,6)) +
  geom_hline(yintercept = 1.3559) +
  labs(y = "Model Estimate") 
```

```{r}
box_renters <- all_cities %>%
  filter(name == "Estimates") %>%
  rename(parameter = `Inputs Parameters`) %>%
  mutate(Significance = ifelse(pvalue <= 0.05, "< 0.05", "not")) %>%
  filter(parameter == "Log Percent Renters") %>%
  filter(Significance == "< 0.05")

renters <- ggplot(box_renters, aes(x = Location, y = value, colour = Significance, shape = Significance)) +
  geom_point() +
  geom_hline(yintercept = 0.5669) +
  theme_few() +
  scale_shape_manual(values = c(15,16),
                     breaks = c("< 0.05", "not"),
                      labels = c("< 0.05", "not")) +
  scale_colour_manual(values = c("#e41a1c", "#4daf4a"),
                      breaks = c("< 0.05", "not"),
                      labels = c("< 0.05", "not")) +
  scale_y_continuous(limits = c(-5,6)) +
  labs(y = "Model Estimate") 
```


```{r}
box_income <- all_cities %>%
  filter(name == "Estimates") %>%
  rename(parameter = `Inputs Parameters`) %>%
  mutate(Significance = ifelse(pvalue <= 0.05, "< 0.05", "not")) %>%
  filter(parameter == "Log Median Income") %>%
  filter(Significance == "< 0.05")

income <- ggplot(box_income, aes(x = Location, y = value, colour = Significance, shape = Significance)) +
  geom_point() +
  theme_few() +
  scale_shape_manual(values = c(15, 17, 16),
                     breaks = c("< 0.05", "not"),
                      labels = c("< 0.05", "not")) +
  scale_colour_manual(values = c("#e41a1c", "#4daf4a"),
                      breaks = c("< 0.05", "not"),
                      labels = c("< 0.05", "not")) +
  labs(y = "Model Estimate") +
  scale_y_continuous(limits = c(-5,6)) +
  theme(legend.position = "bottom")

leg <- get_legend(income)
```

```{r}
box_degrees <- all_cities %>%
  filter(name == "Estimates") %>%
  rename(parameter = `Inputs Parameters`) %>%
  mutate(Significance = ifelse(pvalue <= 0.05, "< 0.05", "not")) %>%
  filter(parameter == "Log Percent Collage Degrees") %>%
  filter(Significance == "< 0.05")

degrees <- ggplot(box_degrees, aes(x = Location, y = value, colour = Significance, shape = Significance)) +
  geom_point() +
  geom_hline(yintercept = 1.7298) +
  theme_few() +
  scale_shape_manual(values = c(15,16),
                     breaks = c("< 0.05", "not"),
                      labels = c("< 0.05", "not")) +
  scale_colour_manual(values = c("#e41a1c", "#4daf4a"),
                      breaks = c("< 0.05", "not"),
                      labels = c("< 0.05", "not")) +
  scale_y_continuous(limits = c(-5,6)) +
  labs(y = "Model Estimate")

parameters <- ggarrange(color, renters, income, degrees,
                        labels = c("POC", "Renters", "Income", "Degrees"),
                        legend = "none")
```

# City affect 
### To compare the cities on the front range (Fort Collins, Greeley, Boulder, Denver, Colorado Springs, and Pueblo), we created a model for the whole front range that predicts ecoli from our four demographic parameters. We also ran the model for each city individually. Form this we can compare the model estimates for each demographic parameter in one city to another and to the front range as a whole. So if a city has a higher estimate for POC than another city, a block group with the same amount of POC would have higher ecoli concentrations in that city than the other. This allows us to see how environmental justice patterns is not consistent throughout the front range. 

### I tried to display the models for each city in different ways. These two tables and two graphs are all the same data, just displayed different.

## Model Output 
### Model estimates for each city and the front range shown as "Global"
```{r, include=TRUE, warning=FALSE}
global <- front_range_m %>%
  pivot_longer(cols = c("Estimates", "Std. Error")) %>%
  filter(name == "Estimates") %>%
  rename(Parameter = `Inputs Parameters`,
         'P-Value' = 2,
         "Model Estimate" = value) %>%
  filter(Parameter != "Intercept") %>%
  mutate(Location = "Global") %>%
  select(-name)

cities_table <- rbind(box_degrees, box_color, box_renters, box_income) %>%
  select(-name, -Significance) %>%
  rename("Model Estimate" = value) %>%
  mutate(pvalue = round(pvalue, 5)) %>%
  arrange(Location) %>%
  select(Location, Parameter=parameter, 4, "P-Value"=pvalue) %>%
  rbind(global) 

kable(cities_table) %>%
  kable_styling(latex_options = "striped")
  
```

```{r}
kable(tibble("City" = c("Fort Collins", "Greeley", "Boulder", "Denver", "Colorado Springs", "Pueblo"),
       "Significant parameters above Front Range" = c("None", "None", "None", "POC, Degrees", "Renters, Income", "POC"),
       "Significant parameters below Front Range" = c("None", "POC", "None", "Income", "None", "Renters"))) %>%
  kable_styling(latex_options = "striped")
```


### I created two graphs that display the front range and city models in different ways. First grouping by demographic parameter and the second grouping by city.


## Compare by demographics 
### The horizontal lines show the model estimates for the whole front range and the symbols show the model estimates for each city. The shapes and color show the estimate's p-value for each city.

```{r, fig.width=8, include = TRUE}
parameters
```

```{r, include = TRUE, fig.height=1}
as_ggplot(leg)
```


```{r}
whole_state <- tibble("Inputs Parameters" = c("Intercept", "Log Percent People of Color", "Log Percent Renters", "Log Median Income", "Log Percent Collage Degrees"),
                         "pvalue" = c(0.833, 0.000005, 0.0218, 0.1866, 0.00378),
                         "Location" = "Global",
                         "name" = "Estimates",
                         "value" = c(-1.1026, 1.3559, 0.5669, -0.761, 1.7298))

ft_graphe <- ft_m %>%
  rbind(whole_state) %>%
  filter(name == "Estimates") %>%
  rename(parameter = `Inputs Parameters`) %>%
  mutate(parameter = case_when(parameter == "Log Percent People of Color" ~ "POC",
                               parameter == "Log Percent Renters" ~ "Renters",
                               parameter == "Log Median Income" ~ "$",
                               parameter == "Log Percent Collage Degrees" ~ "Degrees",
                               parameter == "Intercept" ~ "Int")) %>%
  mutate(Significance = ifelse(pvalue <= 0.05, "< 0.05", "not")) %>%
  filter(Significance == "< 0.05") %>%
  ggplot(aes(x = parameter, y = value, shape = Significance, colour = Location)) +
  geom_point() +
  theme_few() +
  scale_colour_manual(values = c("#74a9cf", "#252525"),
                      breaks = c("Global", "FC"),
                      labels = c("Global", "FC")) +
  scale_y_continuous(limits = c(-4.5,6)) +
  labs(x = "Parameter", y = "Model Estimate")
```

```{r}
denver_graphe <- denver_m %>%
  rbind(whole_state) %>%
  filter(name == "Estimates") %>%
  rename(parameter = `Inputs Parameters`) %>%
  mutate(parameter = case_when(parameter == "Log Percent People of Color" ~ "POC",
                               parameter == "Log Percent Renters" ~ "Renters",
                               parameter == "Log Median Income" ~ "$",
                               parameter == "Log Percent Collage Degrees" ~ "Degrees",
                               parameter == "Intercept" ~ "Int")) %>%
  mutate(Significance = ifelse(pvalue <= 0.05, "< 0.05", "not")) %>%
  filter(Significance == "< 0.05") %>%
  ggplot(aes(x = parameter, y = value, shape = Significance, colour = Location)) +
  geom_point() +
  theme_few() +
  scale_colour_manual(values = c("#252525", "#74a9cf"),
                      breaks = c("Global", "Denver"),
                      labels = c("Global", "Denver")) +
  scale_y_continuous(limits = c(-4.5,6)) +
  labs(x = "Parameter", y = "Model Estimate")
```

```{r}
csprings_graphe <- csprings_m %>%
  rbind(whole_state) %>%
  filter(name == "Estimates") %>%
  filter(1 != "Intercept") %>%
  rename(parameter = `Inputs Parameters`) %>%
  filter(parameter != "Intercept") %>%
  mutate(parameter = case_when(parameter == "Log Percent People of Color" ~ "POC",
                               parameter == "Log Percent Renters" ~ "Renters",
                               parameter == "Log Median Income" ~ "$",
                               parameter == "Log Percent Collage Degrees" ~ "Degrees")) %>%
  mutate(Significance = ifelse(pvalue <= 0.05, "< 0.05", "not")) %>%
  filter(Significance == "< 0.05") %>%
  ggplot(aes(x = parameter, y = value, shape = Significance, colour = Location)) +
  geom_point() +
  theme_few() +
  scale_colour_manual(values = c("#252525", "#74a9cf"),
                      breaks = c("Global", "CSprings"),
                      labels = c("Global", "CSprings")) +
  scale_y_continuous(limits = c(-4.5,6)) +
  labs(x = "Parameter", y = "Model Estimate")
```

```{r}
pueblo_graphe <- pueblo_m %>%
  rbind(whole_state) %>%
  filter(name == "Estimates") %>%
  rename(parameter = `Inputs Parameters`) %>%
  mutate(parameter = case_when(parameter == "Log Percent People of Color" ~ "POC",
                               parameter == "Log Percent Renters" ~ "Renters",
                               parameter == "Log Median Income" ~ "$",
                               parameter == "Log Percent Collage Degrees" ~ "Degrees",
                               parameter == "Intercept" ~ "Int")) %>%
  mutate(Significance = ifelse(pvalue <= 0.05, "< 0.05", "not")) %>% 
  filter(Significance == "< 0.05") %>%
  ggplot(aes(x = parameter, y = value, shape = Significance, colour = Location)) +
  geom_point() +
  theme_few() +
  scale_colour_manual(values = c("#74a9cf", "#252525"),
                      breaks = c("Global", "Pueblo"),
                      labels = c("Global", "City")) +
  scale_y_continuous(limits = c(-4.5,6)) +
  theme(legend.position = "bottom") +
  labs(x = "Parameter", y = "Model Estimate")

city_leged <- get_legend(pueblo_graphe)
```


```{r}
cities <- ggarrange(ft_graphe, denver_graphe, csprings_graphe, pueblo_graphe,
                        labels = c("FC", "Denver", "C Spgs", "Pueblo"),
                        legend = "none")
```

## Compare by city
### Here we are showing the same data as above but grouped by city. Int is the model's intercept estimates and $ is annual median income.

```{r, include=TRUE, warning=FALSE, fig.width=8}
cities
```

```{r, include=TRUE, fig.height=1}
as_ggplot(city_leged)
```

# Ecoli sites in cities
### This is just a table with the number of ecoli sites in each metro area and the number of these sites on a "major" stream in that city. I considered the streams in the next section as major streams.

```{r}
city_site_info  <- tibble("City" = c("Fort Collins", "Greeley", "Boulder", "Denver", "Colorado Springs", "Pueblo"),
       "Number" = c(14, 21, 19, 58, 41, 18),
       "On_Major_River" = c(8, 12, 10, 20, 28, 16)) %>%
  mutate(percent = (On_Major_River/Number)*100)

```

```{r, include=TRUE, warning=FALSE}
kable(city_site_info)
```


# Landcover and WWTP 
### To look at how landcover and WWTP impact ecoli concentration, we plotted the landcover of the watershed longitudinally as you move down the main stem of each stream. The vertical lines represent the locations of wastewater treatment plants (WWTP). 

### These plots show that streams on the front range have similar landcovers patterns. Their headwater are dominated by forests and as they inter the front range the percent of forest declines and the percent of shrub land, urbans area, and agriculture increases. We also see that after urban areas (and WWTPs) ecoli concentrations spike. Also as the percent of agriculture increases, ecoli concentrations generally rise.

## Poudre 
### Fort Collins Metro Areas  
```{r, include=TRUE, warning=FALSE}
poudre <- read_rds("/Users/spencerrhea/Google Drive/EJ_paper /data_out/poudre.rds")

ggplot(poudre[["full"]], aes(x = pathlength, y = value, colour = name)) +
  geom_vline(xintercept = poudre[["wwtp"]], color = "gray") +
  geom_line() +
  geom_point(aes(x = pathlength, y = mean)) +
  scale_color_manual(values = c("#252525", "#e7298a", "#66a61e", "#7570b3", "#d95f02", "#1b9e77"), 
                     breaks = c("herbaceous", "shrub", "forest", "wetland", "agriculture", "developed"),
                     name = "Landcover") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*50,  name = "E.coli (#/ml)")) +
  theme_few() +
  labs(x = "Length on River", y = "Watershed landcover (%)")
```

## Big Thompson 
### Fort Collins Metro Areas
```{r, include=TRUE, warning=FALSE}
bt_full <- read_rds("/Users/spencerrhea/Google Drive/EJ_paper /data_out/big_thompson.rds")

ggplot(bt_full[["full"]], aes(x = pathlength, y = value, colour = name)) +
  geom_vline(xintercept = bt_full[["wwtp"]], color = "gray") +
  geom_line() +
  geom_point(aes(x = pathlength, y = mean)) +
  scale_color_manual(values = c("#252525", "#e7298a", "#66a61e", "#7570b3", "#d95f02", "#1b9e77"), 
                     breaks = c("herbaceous", "shrub", "forest", "wetland", "agriculture", "developed"),
                     name = "Landcover") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*50,  name = "E.coli (#/ml)")) +
  theme_few() +
  labs(x = "Length on River", y = "Watershed landcover (%)")
```

## South Platte 
### Denver Metro Area
```{r, include=TRUE, warning=FALSE}
south_platte <- read_rds("/Users/spencerrhea/Google Drive/EJ_paper /data_out/south_platte.rds")

sp_sweage_codes <- c(281.122, 281.122, 188.124, 188.124, 178.222, 178.222, 178.222, 178.222, 178.222, 178.222, 157.964, 157.964, 157.964, 157.964, 157.964, 157.964)

ggplot(south_platte, aes(x = pathlength, y = value, colour = name)) +
  geom_vline(xintercept = sp_sweage_codes, color = "gray") +
  geom_line() +
  geom_point(aes(x = pathlength, y = mean)) +
  scale_color_manual(values = c("#252525", "#e7298a", "#66a61e", "#7570b3", "#d95f02", "#1b9e77"), 
                     breaks = c("herbaceous", "shrub", "forest", "wetland", "agriculture", "developed"),
                     name = "Landcover") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*50,  name = "E.coli (#/ml)")) +
  theme_few() +
  labs(x = "Length on River", y = "Watershed landcover (%)")
```

## Monument Creek 
### Colorado Springs Metro Area
```{r, include=TRUE, warning=FALSE}
monument_creek <- read_rds("/Users/spencerrhea/Google Drive/EJ_paper /data_out/monument_creek.rds")

ggplot(monument_creek[["full"]], aes(x = pathlength, y = value, colour = name)) +
  geom_vline(xintercept = monument_creek[["wwtp"]], color = "gray") +
  geom_line() +
  geom_point(aes(x = pathlength, y = mean)) +
  scale_color_manual(values = c("#252525", "#e7298a", "#66a61e", "#7570b3", "#d95f02", "#1b9e77"), 
                     breaks = c("herbaceous", "shrub", "forest", "wetland", "agriculture", "developed"),
                     name = "Landcover") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*50,  name = "E.coli (#/ml)")) +
  theme_few() +
  labs(x = "Length on River", y = "Watershed landcover (%)")
```


## Fountain Creek
### Colorado Springs and Pueblo Metro Areas
```{r, include=TRUE, warning=FALSE}
fountain_full <- read_rds("/Users/spencerrhea/Google Drive/EJ_paper /data_out/fountain.rds")

ggplot(fountain_full[["full"]], aes(x = pathlength, y = value, colour = name)) +
  geom_vline(xintercept = fountain_full[["wwtp"]], color = "gray") +
  geom_line() +
  geom_point(aes(x = pathlength, y = mean)) +
  scale_color_manual(values = c("#252525", "#e7298a", "#66a61e", "#7570b3", "#d95f02", "#1b9e77"), 
                     breaks = c("herbaceous", "shrub", "forest", "wetland", "agriculture", "developed"),
                     name = "Landcover") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*50,  name = "E.coli (#/ml)")) +
  theme_few() +
  labs(x = "Length on River", y = "Watershed landcover (%)")
```


## Arkansas 
### Pueblo Metro Areas
### This river may not be comparable to the other because of its size. It also has a large reservoir just above Pueblo.
```{r, include=TRUE, warning=FALSE}
arkansas_full <- read_rds("/Users/spencerrhea/Google Drive/EJ_paper /data_out/arkansas.rds")

ggplot(arkansas_full[["full"]], aes(x = pathlength, y = value, colour = name)) +
  geom_vline(xintercept = arkansas_full[["wwtp"]], color = "gray") +
  geom_line() +
  geom_point(aes(x = pathlength, y = mean)) +
  scale_color_manual(values = c("#252525", "#e7298a", "#66a61e", "#7570b3", "#d95f02", "#1b9e77"), 
                     breaks = c("herbaceous", "shrub", "forest", "wetland", "agriculture", "developed"),
                     name = "Landcover") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*50,  name = "E.coli (#/ml)")) +
  theme_few() +
  labs(x = "Length on River", y = "Watershed landcover (%)")
```
